---
output:
  word_document:
    reference_docx: myref.docx
bibliography: MyBib.bib
csl: Evolution.csl

---
```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE)

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')


```




**Early-life conditions, telomere dynamics and survival: insights from a long-term island bird study**

Lewis G. Spurgin^1,2^,  Kat Bebbington^1^, Eleanor A. Fairfield^1^, Martijn Hammers^3^, Jan Komdeur^3^, Terry Burke^4^, Hannah, L. Dugdale^3,4^, and David S. Richardson^1,5,^.

1. School of Biological Sciences, University of East Anglia, Norwich Research Park, NR4 7TJ, United Kingdom
2. Department of Zoology, Edward Grey Institute, University of Oxford, Oxford, UK
3. Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands
4. Department of Animal and Plant Sciences, University of Sheffield, Sheffield, UK
5. Nature Seychelles, Roche Caiman, Mahé, Republic of Seychelles

**Correspondence:** Lewis Spurgin: lewisspurgin@gmail.com; David Richardson: david.richardson@uea.ac.uk

**Running head:** Early-life telomeres and survival

###Abstract
1. Understanding the short and long term costs of individual early-life experiences is fundamental to understanding life-history evolution. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, it is possible that telomere dynamics may help us quantify individual variation in early-life costs, and enhance our understanding of how poor conditions in early life are related to later-life survival. 

2. We tested how telomere dynamics are related to spatio-temporal variation in early-life conditions and later-life survival in the Seychelles warbler (*Acrocephalus sechellensis*), using a large dataset of `r nrow(juv)` juveniles from multiple cohorts spanning `r max(juv$LayYear)-min(juv$LayYear)` years. 

3. We found that at the population level, telomere length decreases with cross-sectional and longitudinal measures of age, and that telomere length decreases most rapidly early in life. However, at the individual level we found only a very weak relationship between juvenile and adult telomere length, and that significant reductions in telomere length can only be detected within the first few months of life - after this, longitudinal changes in telomere length were not significantly different from zero. 

4. Juvenile telomere length varied markedly among cohorts, and birds born in years with higher populaiton density had longer telomeres. Within cohorts, we found no evidence that the early-life social environment or ecological conditions were related to telomere length or rates of shortening. We also found no relationship between early-life telomere dynamics and survival later in life. 

5. Our data suggest that telomeres are subject to strong population-level variation, but suggest that the resolution of telomeres as a useful biomarker of individual-level cost is likely to be age-dependent, and in many cases limited.

**Keywords:** Life-history; Seychelles warbler; Senescence; Survival

**Data archival location:** This manuscript was written in R Markdown (http://rmarkdown.rstudio.com/). All data and scripts required to reproduce the manuscript, figures and analyses will be made available on GitHub.

PAGEBREAK

###Introduction
Exposure to favourable environmental conditions during development and growth can confer fitness advantages later in life [so called 'silver spoon effects'; @Grafen1988; @Monaghan2008]. Adult fitness can be affected by a range of early-life experiences, including the quality of the habitat and available resources [@Madsen2000; @VandePol2006; @Hayward2013], population density [@Nussey2007;@Douhard2013] and natural or anthropogenic environmental disturbance [@Reid2003; @Cartwright2014]. Understanding the causes and consequences of these early-life experiences is key to understanding many ecological and evolutionary  processes, including  patterns of natural and sexual selection,  population growth rates and even local extinction [@Coulson2001; @Roach2014]. Understanding silver spoon effects is therefore of central interest to ecologists, evolutionary biologists and conservationists.

While it is clear that silver spoon effects can occur, we have little understanding of how and under what conditions early-life environments will affect adult fitness. We expect the later-life consequences of a good or poor start to vary among individuals, populations and species because the phenotypic consequences of an adverse environment can depend on an individual's initial condition, and genetic or epigenetic makeup [@Hoffman2000; @Richards2006]. Moreover, it is not always possible to fully quantify what constitutes a good or bad environment, and any 'hidden' environmental variation may obscure relationships between the early-life conditions that are measured and the resulting adult phenotypes. Indeed, the pervasiveness of silver spoon effects varies between species [@Drummond2011], cohorts [@Reid2003] and sexes [@Wilkin2009]. We therefore need to understand, or at least be able to measure, how the environment affects individuals differentially within a population in order to elucidate the later-life consequences of early-life experience.

Telomeres - the protective caps on the ends of chromosomes - may provide a solution to this problem. Telomeres shorten with age [@Monaghan2006], and in response to oxidative stress, which can be elevated  due to environmental factors [@VonZglinicki2002]. When telomeres become critically short cells senesce [@Campisi2003], and the accumulation of these cells can result in organismal senescence and death  [@Wong2003]. This association between senescence and telomere length has inspired a great deal of recent research into telomere evolutionary ecology [reviewed in @Horn2010; @Haussmann2010; @Monaghan2014]. While there is little direct evidence that the relationship between telomere dynamics and survival is causal [@Simons2015], there is mounting evidence that telomeres can act as biomarkers of cost in wild populations, providing a signature of the ecological stress that has been experienced and is otherwise difficult to detect [@Monaghan2014; @Schultner2014; @Asghar2015].

There is evidence from a range of taxa that the greatest rate of telomere shortening occurs in early life [e.g. @Frenck1998; @Haussmann2003], and that the extent of this shortening is influenced by the conditions experienced during that period [@Price2013; @Monaghan2014; @Nettle2015; @Reichert2015]. Early-life telomere dynamics have been associated with both short-term and late-life survival [@Heidinger2012; @Boonekamp2014], and with other parameters such as cognition [@Nettle2015a]. However, few studies have simultaneously analysed how telomere dynamics, early-life conditions and late-life survival are all related in a natural setting. Moreover, how early-life telomere dynamics vary over spatial and temporal scales is poorly understood.

The longitudinal study (since 1986) of the Seychelles warbler (*Acrocephalus sechellensis*) population on Cousin Island provides an excellent system for studying senescence in the wild [reviewed in @Hammers2015]. Due to the isolated nature of the study population and intensive field monitoring, we have unusually comprehensive survival data and tissue samples spanning many years (see Methods, below). Ecological conditions and warbler population density on Cousin vary across space and time due to weather-induced changes in foliage cover and food availability [@VandeCrommenacker2011a]. Seychelles warblers remain on their natal territories for at least six months, and variation in the oxidative stress experienced by individuals is associated with natal  territory quality [@VandeCrommenacker2011a]. However, neither early-life nor adult survival appear to be associated with natal territory quality or natal local density [@Brouwer2006; @Hammers2013]. Facultative cooperative breeding occurs in the Seychelles warbler [@Komdeur1994a; @Richardson2003a], and the presence of helping subordinates (but not non-helping subordinates) in the natal territory is associated with increased survival later in life [@Brouwer2012]. Lastly, we have an established protocol for assessing absolute telomere length in this species [@Barrett2012], and telomere length predicts survival independently of age in adult Seychelles warblers, suggesting that telomeres act as a biomarker of cost [@Barrett2013a]. Thus, we have an excellent system in which to assess the costs of different social and environmental conditions experienced early in life, and to assess the later-life consequences of early-life conditions.

In this study we examine how telomeres are related to early-life environmental variation and short and long-term survival in the Seychelles warbler. Because telomere dynamics are expected to reflect individual-level variation in the costs of early-life experiences, they may allow a more sensitive analysis of the effects of early-life environmental variation than would be possible with a direct comparison of how survival is affected by the early-life environment. With this in mind, we first conduct an exploratory analysis of how the environmental and social factors experienced in early life affect telomere dynamics. We then test the hypothesis that longer telomeres and lower rates of telomere shortening in early life are associated with greater survival.


###Methods

#####Study species and sampling
The Seychelles warbler is a small (~15 g), insectivorous passerine bird with a mean life expectancy of 5.5 years at fledging [@Hammers2013]. The population of *ca*. 320 birds on Cousin Island (04'20'S, 55'40'E) has been intensively studied since 1986 [@Richardson2003b; @Spurgin2014a]. This species has two breeding seasons, running from June-August (main breeding season) and December-February (minor breeding season), when the breeding females on each of the *ca*. 115 territories lay one or, rarely, two or three eggs [@Komdeur1991]. As a result of this low reproductive output, combined with higher mortality in first-year birds [39%; @Brouwer2006], cohort sizes in the Seychelles warbler are small (<50).

Individuals are usually ringed in their first year of life, and so are of known age. They are then followed throughout their lives, and as they are non-migratory endemics naturally confined to the island [@Komdeur2004a], a biannual census of birds on Cousin during each breeding season gives accurate measures of local density, social status (e.g. breeding male/female, helping subordinate, non-helping subordinate) and individual survival [@VandeCrommenacker2011; @Barrett2013a]. The isolated nature of the Cousin population is a key advantage of the system for analyses involving survival, which in other systems are often confounded by emigration [see @Ergon2014 for a recent discussion]. Full details of catching and monitoring methods can be found in Brouwer *et al.* [-@Brouwer2012].

Seychelles warblers are highly territorial and all territories are mapped during the breeding seasons using detailed observational data of foraging and territorial defence behaviour, and surveyed for territory quality [@Richardson2003b]. Territory quality is calculated based on territory size, foliage cover and insect abundance [@Komdeur1992], and territory quality estimates obtained across years are averaged to obtain a single value for each territory [@Hammers2013]. Cousin is subject to intra- and inter-annual variation in rainfall and food availability, and such island-wide temporal variation may override the effects of individual territory quality. As an estimate of seasonal variation in food availability, we calculated an index of the number of insects across the entire island during each breeding season. This index is calculated as the mean number of insects found per unit leaf area over all surveys carried out on the island in a breeding season.

Each time a bird is caught on Cousin body mass and tarsus length are measured (to the nearest 0.1g and 0.1mm, respectively). Using information on eye colour [@Komdeur1991a] and previous captures [@Richardson2003b], we grouped birds into three age categories: one month old (birds still in the nest), 6 months old (fledglings with light grey eyes) and 10 months old (subadults with light brown eyes). A blood sample (*ca* 25 $\mu$l) is taken from each bird captured via brachial venipuncture, and stored at room temperature in 1 ml of absolute ethanol in a 1.5 ml screw-cap microfuge tube.


#####Molecular methods
For each sample, genomic DNA was extracted from a ~2 mm^2^ flake of preserved blood using the DNeasy Blood and Tissue Kit (Qiagen), following the manufacturer's protocol, with the modification of overnight lysis at 37^o^C and a final DNA elution volume of 80 $\mu$l.  Sex was determined using the PCR-based method outlined by Griffiths *et al.* [-@Griffiths1998]. Prior to telomere analysis, DNA concentration and purity were quantified using a NanoDrop 8000 Spectrophotometer (ThermoScientific), and the following thresholds were applied before samples were included for further analysis: i) DNA concentration must be at least 15 ng $\mu$l^-1^ (based on a mean of three measurements), ii) the 260/280 ratio has to be between 1.8 and 2 and, iii) the 260/230 ratio should be higher than 1.8. DNA integrity was further validated by visualization with ethidium bromide after electrophoresis on a 1.2% agarose gel, and all samples with evidence of DNA degradation were re-extracted or excluded.

All DNA extractions that passed the above criteria were diluted to 3.3 ng $\mu$l^-1^ before telomere measurement. We measured telomere length for all samples using a quantitative PCR (qPCR) assay of telomeres and a GAPDH control gene, using the molecular methods outlined by Barrett *et al.* [-@Barrett2012], with two amendments. A change in batch of SYBR green forced us to raise the annealing temperature of the telomere reaction from 58^o^C to 61^o^C for the majority of samples. However, this did not affect final telomere length values (see below). Second, to avoid bacth effects, we randomised samples among plates. We used the program LinRegPCR [@Ruijter2009] to correct for baseline fluorescence and calculate efficiencies and Cq values for each sample replicate. Averaging of technical repeats was carried out using custom-made R scripts (available as supplementary material), excluding samples with Cq values differing by >0.5. We then calculated relative telomere length (RTL) for each sample using equation 1 in Pfaffi *et al* [@Pfaffi2001]. We chose to use RTL rather than continuing with the previously used method for calculating absolute telomere length [@Barrett2012], as i) using RTL enabled us to run more samples per plate (as an oligo standard is not required), ii) the RTL method was less susceptible to batch effects (Appendix 1), and iii) very few other studies have calculated absolute telomere length, and our experience suggests that cross-species comparisons are unlikely to be reliable. Inter-plate repeatability of final telomere lengths was assessed using the R package rptR [@Schielzeth2011], and for all subsequent analyses we used mean telomere length per sample where we had repeats.

For a subset of first-year birds for which we had longitudinal data, with an additional sample taken as an adult. For these individuals we calculated the within-individual change in RTL by subtracting adult RTL from early-life RTL (hereafter $\Delta$RTL). Negative values of $\Delta$RTL reflect telomere shortening, while positive values reflect telomere lengthening. 


#####Statistical analyses
We performed all analyses using R version 3.0.1 [@RDevelopmentCoreTeam2011]. Because we are using a different measure of telomere length to the previous study of adult telomere dynamics in this species [@Barrett2013], as well as including many additional samples, we first re-tested, using all available samples, whether telomere length decreases with age and $\Delta$age [a longitudinal measure based on within-subsject centring; @VandePol2009] in the Seychelles warbler, using linear mixed models, with individual ID as a random effect. For the analysis of $\Delta$age we we only included birds who had been sampled at least twice. We then explored how telomere length varied within the first year of life (telomere length vs age in months) and among cohorts (telomere length vs cohort ID), again using linear mixed models with individual ID as a random effect.

To assess our resolution for detecting telomere shortening, we tested, for each age class, whether $\Delta$RTL was significantly lower than zero, using one-sample t-tests. We then tested how $\Delta$RTL was related to the time interval between sampling events, using linear models. A significantly negative slope indicates that telomere shortening occurs linearly over time, while the time point at which $\Delta$RTL is significantly lower than zero can be considered the time period over which we have the resolution to detect reductions in telomere length. 

We used general linear mixed models with model averaging to explore how spatial variation in early-life environmental and social conditions influenced telomere length within cohorts. Model averaging was carried out using the MuMIn package (version 1.10.5) in R [@Barton2012]. We first created a full model containing the following explanatory variables: age (in months), tarsus length, sex, territory quality, season (major or minor) and the number of helping and non-helping subordinate birds present in the natal territory. We also included interaction terms between age and all the other variables as telomere dynamics are expected to vary within the first year of life [@Heidinger2012]. As random effects we included cohort, territory ID and qPCR plate ID. Model selection was then performed and a top model set defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(Loss)`) we repeated the above analyses of telomere dynamics, replacing telomere length with telomere loss as the response variable, and excluding the plate ID random effect (as each measurement of telomere loss was based on two or more measurements, and so run on multiple plates).

We used generalised linear mixed models to test whether telomere length/loss were related to survival to adulthood. Models were set up with a binomial error structure, and included RTL/$\Delta$RTL and juvenile age as explanatory variable, along with the same random effects as for the model averaging (see above). We then used Cox regression, implemented in the survival package in R [@Therneau2011], to test whether long-term survival was related to individual-level telomere dynamics. For the cross-sectional data we used lifespan (in years) as survival time, while for the longitudinal data we used remaining lifespan from the second sampling event. We ran these models with cohort as a frailty term to exclude cohort effects, and only included cohorts in which at least 70 percent of individuals had died (all years up to and including 2007).

###Results
We measured telomere lengths using a total of `r nrow(dd)` unique samples from juvenile and adult Seychelles warblers. Of these, `r nrow(juv)` samples were taken cross-sectionally from birds caught within their first year of life, and of these juvenile birds, we had additional longitudinal samples for `r nrow(Loss)`. Efficiencies (mean $\pm$) for our telomere and GAPDH reactions were (`r round2(mean(dd$EffTelomere))` $\pm$ `r round2(sd(dd$EffTelomere))`) and (`r round2(mean(dd$EffGAPDH))` $\pm$ `r round2(sd(dd$EffGAPDH))`) respectively. Inter-plate repeatability of RTL, based on 392 samples measured at least twice, was 0.78 (CI = 0.73-0.82).

#####Early-life RTL
Considering the entire Seychelles warbler lifespan, RTL decreased with age (LMM: t = `r round2(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=dd,random=~1|BirdID))$tTable[2,5])`), although there were clear increases in cross-sectional telomere length after one year of age and at several points later in life (Fig. 1A). RTL also decreased with age within individuals (i.e. with $\Delta$age; t = `r round2(summary(lme(RTL~DeltaAge,data=subset(dd,DeltaAge!=0),random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~DeltaAge,data=subset(dd,DeltaAge!=0),random=~1|BirdID))$tTable[2,5])`; Fig. 1B). RTL also decreased with age within the first year of life (i.e. excluding adult samples; t = `r round2(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID))$tTable[2,4])`; P `r round3(summary(lme(RTL~Agemonths,data=juv,random=~1|BirdID))$tTable[2,5])`). 

Early-life RTL varied significantly among cohorts (Likelihood ratio test: F = `r round2(anova(lme(RTL~factor(FieldPeriodID),data=juv,random=~1|BirdID))[2,3])`; P `r round3(anova(lme(RTL~Agemonths,data=juv,random=~1|BirdID))[2,4])`). Cohort-level variation in juvenile RTL was positively related to adult population density (LMM controlling for age and cohort: t = `r round2(summary(lme(RTL~Agemonths+Density,data=juv,random=~1|FieldPeriodID))$tTable[3,4])`; P `r round3(summary(lme(RTL~Agemonths+Density,data=juv,random=~1|FieldPeriodID))$tTable[3,5])`; Figs 2B,2C). Variation in cohort-level RTL was not related temporal variation in  food availability (t = `r round2(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,4])`; P `r round3(summary(lme(RTL~Agemonths+Insect,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,5])`). RTL did not differ in birds born in minor compared to major seasons (t = `r round2(summary(lme(RTL~Agemonths+Season,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,4])`; P `r round3(summary(lme(RTL~Agemonths+Season,data=juv,random=~1|FieldPeriodID,na.action=na.exclude))$tTable[3,5])`; Fig. 2A, 2C).

Within cohorts, the top model explaining variation in early-life RTL contained age, sex, tarsus length, and the interaction between age and tarsus length (Table S1). The top model was much better supported than a null model ($\Delta$AICc = `r round2(AICc(lmer(RTL~1+(1|FieldPeriodID)+(1|TerritoryID),data=juv,REML = FALSE))-AICc(lmer(RTL~Agemonths+Sex+Tarsus+Agemonths*Tarsus+(1|FieldPeriodID)+(1|TerritoryID),data=juv,REML = FALSE)))`), and with the exception of the sex and tarsus effects, model-averaged confidence intervals did not overlap zero (Fig. 3A). Tarsus length was negatively related to RTL in nestlings, unrelated to RTL in fledglings, and positively related to RTL in subadults (Fig. 3B). Males had slightly longer telomeres than females, particularly at the fledgling and subadult stages, although this effect was weak  (Fig. 3C).

#####Early-life telomere shortening
Longitudinal data showed that early-life RTL was positively, but weakly, related to adult RTL (R^2^ = `r round3(summary(lm(RTL1~RTL+Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$coefficients[2,3])`; P = `r round3(summary(lm(RTL1~RTL+Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4A). $\Delta$RTL was significantly lower than zero in nestlings (t = `r round2(t.test(subset(Loss,Agemonths == 1)$Loss)$statistic)`; P `r round3(t.test(subset(Loss,Agemonths == 1)$Loss)$p.value)`), but not in fledglings (t = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$p.value)`) or subadults (t = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$p.value)`). In line with this, $\Delta$RTL increased with age in early life (R^2^ = `r round2(summary(lm(Loss~Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(Loss~Agemonths,data=Loss))$coefficients[2,3])`; P `r round3(summary(lm(Loss~Agemonths,data = Loss))$coefficients[2,4])`), suggesting that telomere shortening was highest shortly after the nestling phase. In nestlings, $\Delta$RTL  did not vary with the time between sampling events (R^2^ `r round3(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths ==1)))$r.squared)`; t = `r round2(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths ==1)))$coefficients[2,3])`; P = `r round3(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths ==1)))$coefficients[2,4])`). However, in fledglings and subadults $\Delta$RTL decreased with the length of time between sampling events (R^2^ = `r round2(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths !=1)))$r.squared)`; t = `r round2(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths !=1)))$coefficients[2,3])`; P = `r round3(summary(lm(Loss~TimeDiff,data=subset(Loss,Agemonths !=1)))$coefficients[2,4])`; Fig. 4B, Fig. S1). In both fledglings and subadults, $\Delta$RTL did not decrease below zero until ~3 years between sampling events was allowed, and confidence limits overlapped zero regardless of the amount of time between sampling events (Fig 3B). Together, these results suggest that the majority of observable telomere shortening in Seychelles warblers occurs in the first six months of life. We therefore restricted the remaining analyses of $\Delta$RTL to birds sampled as nestlings (N = `r nrow(subset(Loss,Agemonths == 1))`. 

$\Delta$RTL did not differ significantly among cohorts of nestlings (F = `r round2(anova(lm(Loss~factor(FieldPeriodID),data = subset(Loss,Agemonths == 1)))[1,4])`; P `r round3(anova(lm(Loss~factor(FieldPeriodID),data = subset(Loss,Agemonths == 1)))[1,5])`), although power for this analysis was limited as longitudinal sample sizes within cohorts were low. Model selection revealed that the top model explaining $\Delta$RTL in nestlings contained sex and tarsus length (Table S2). This model was a better fit than the null model ($\Delta$AICc = `r round2(AICc(lmer(Loss~1+(1|FieldPeriodID)+(1|TerritoryID),data=subset(Loss,Agemonths == 1),REML = FALSE))-AICc(lmer(Loss~Sex+Tarsus+(1|FieldPeriodID)+(1|TerritoryID),data=subset(Loss,Agemonths == 1),REML = FALSE)))`), although confidence limits for both the sex and tarsus effects overlapped zero (Fig. 5A). $\Delta$RTL was negatively related to tarsus length, and higher in females, indicating higher rates of telomere loss in smaller nestlings, and in males.


#####Early-life telomere dynamics and survival

Early-life RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,4])`), with no interaction effect between RTL and juvenile age on survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~RTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,4])`). To test whether cohort-level effects on RTL influenced survival we calculated cohort-level RTL (i.e. the median RTL for each cohort) and added this term to a GLMM, and found that cofort-level RTL was not related to survival to adulthood (estimate = `r round2(summary(glmer(SurvivedNext~cohortTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glmer(SurvivedNext~cohortTL+Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[2,4])`), with no interaction with juvenile age (estimate = `r round2(summary(glmer(SurvivedNext~cohortTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glmer(SurvivedNext~cohortTL*Agemonths+(1|TerritoryID)+(1|FieldPeriodID),data=juv13,family = 'binomial'))$coefficients[4,4])`). Using the longitudinal data, we found that $\Delta$RTL had no effect on survival to the year after the second sampling event (estimate = `r round2(summary(glm(SurvivedNext2~Loss,data=subset(subset(Loss,Agemonths == 1),FieldPeriodID<2013),family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext2~RTL+Agemonths,data=subset(subset(Loss,Agemonths == 1),LayYear<2013),family = 'binomial'))$coefficients[2,4])`).

Cox regression showed no effect of individual-level or cohort-level RTL on survival later in life (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL+Agemonths+frailty(LayYear))))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL+Agemonths+frailty(LayYear))))$coefficients[1,6])`; cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Agemonths+frailty(LayYear))))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Agemonths+frailty(LayYear))))$coefficients[1,6])`), and no interaction between RTL and age on survival (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL*Agemonths+frailty(LayYear))))$coefficients[4,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~RTL*Agemonths+frailty(LayYear))))$coefficients[4,6])`; cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL*Agemonths+frailty(LayYear))))$coefficients[4,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL*Agemonths+frailty(LayYear))))$coefficients[4,6])`). $\Delta$RTL was not related to longer-term survival (estimate `r round3(summary(coxph(with(subset(subset(Loss,LayYear<2008),Agemonths == 1),Surv(RemainingLife2+1,Died)~Loss)))$coefficients[1,1])`, P = `r round3(summary(coxph(with(subset(subset(Loss,LayYear<2008),Agemonths ==1 ),Surv(RemainingLife2+1,Died)~Loss)))$coefficients[1,5])`).


###Discussion
Here we use one of the largest datasets to date to assess the relationships between early-life conditions, telomere length and survival in a closed population of Seychelles warblers. We found that, as in other studies, telomeres are lost at the greatest rate  early in life. We found that telomere dynamics are subject to strong cohort effects, and that temporal variation in RTL is population density and season. Within cohorts, we found no evidence that the ecological or social environment influenced telomere dynamics, and we found no evidence that early-life telomere length is related to juvenile or later-life survival.

Our study adds the the now substantial body of literature showing that telomere length decreases with age, and that this decrease is most rapid in early life [e.g. @Frenck1998; @Haussmann2003; @Heidinger2012; @Fairlie2015]. Our finding that tarsus length was related to juvenile RTL in an age-dependent manner (Fig. 3) most likely represents an additional age effect - in passerine birds tarsus length is correlated with age during the nestling stage, but not after fledging [@Ricklefs1976]. Telomere loss is most rapid early in life due to ongoing growth [@Frenck1998], and a negative correlation between RTL and body size in early life is therefore expected [see also @Ringsby2015]. However, despite our clear finding that population-level RTL decreases with age throughout the warbler lifespan (Fig. 1), we found that at the individual-level, juvenile and adult telomere length were weakly correlated, and that telomere shortening can only be detected very early in life (Fig. 4). This is in contrast to other avian studies where birds were reared in laboratory [@Heidinger2012] or nestbox [@Boonekamp2014] conditions, in which juvenile and adult telomere length were highly correlated, and individual-level telomere shortening clearly occurs throughout the juvenile period, and into adulthood. However, in accordance with our study, a recent study on Soay sheep (*Ovis aries*) found a weak correlation between juvenile and adult RTL, and a highly complex pattern of telomere dynamics across the entire lifespan [@Fairlie2015]. Indeed, the complex relationship between RTL and age in our study, with periods of early-life telomere shortening followed by selective disappearance and/or apparent telomere lengthening as individuals age is strikingly similar to that reported by Fairlie *et al.* [-@Fairlie2015]. It is too soon to speculate as to what may be generating this variation among species, but one possibility is that environmental variability itself affects the resolution with which we can observe telomere shortening in natural populations [see also @Fairlie 2015]. As early-life telomere dynamics are investigated in more systems, the causes of this variation should become clear.

Perhaps the clearest result from our study is that RTL varies among cohorts. Very few studies have shown that temporal variation in telomere dynamics occurs in natural populations, and these have been limited in the number of seasons [@Mizutani2013; @Watson2015; @Fairlie2015]. The long-term Seychelles warbler dataset has allowed us to show that temporal variation in telomere dynamics can occur over substantial time periods. Moroevoer, becuase of the larger number of cohorts we have been able to show statistically that that temporal variation in RTL is related to variation in the environment (Fig. 1C). We may expect a negative relationship between population density and RTL, as increased competition for resources is known to effect juvenile telomere length. However, We suggest that population density acts as an indicator of overall island 'quality' for Seychelles warblers. This can clearly be seen from data from the 1980s and 1990s, when adult population size on Cousin increased as habitat quality improved [@Komdeur2005]. Population size has fluctuated around 300 since the late 1990s (Fig. 2B), and captures this more accurately than measures of insect abundance or territory quality. 

Within cohorts, we found no evidence that any of the social or environmental variables we measured affect juvenile RTL. This is surprising, as a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival. Possible explanations for this finding include i) temporal variation in RTL in our data masks our ability to detect spatial trends, ii) within-cohort variation in RTL is explained by a variable that we have not included in our analyses, or iii) that effects are generally weak and levels of noise in our telomere measurements and/or ecological data preclude detection of significant effects. The first of these explanations seems the least likely, as within-cohort variation in RTL greatly exceeds among-cohort variation (Fig. 1C).

While the relationship between mortality and telomeres in adults has been established for some time [@Cawthon2003], only recently has the link between later-life survival and early-life telomere dynamics been studied. In captive zebra finches, juvenile telomere length predicts late-life survival [@Heidinger2012], and in wild bird populations survival to the nestling phase [@Watson2015], and survival to adulthood [@Boonekamp2014], have been linked with early-life telomere dynamics. 

There are numerous avenues for future research into telomere evolutionary ecology in this system and others. If our findings are replicated in other systems, and population-level variation in early-life telomere dynamics is common in nature, this has potential consequences for our understanding of telomere dynamics in natural populations. In particular, our findings suggest that the telomere dynamics of a population at a given point in time represent a snapshot of a temporally varying process. More research is now needed within and across multiple cohorts and populations to better understand how how and why population-level telomere dynamics vary over space and time. Another question to be addressed is the extent to which early-life RTL reflects inheritance [e.g. @Becker2015; @Asghar2014]. This is important, as inheritance may capture a large amount of unexplained variation in early-life RTL within our dataset. Future research on the Seychelles warbler will examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence in natural populations. FInally, here we have considered survival, but telomere length and shortening in early life may also be linked to other components of fitness [reviewed in @Monaghan2014]. Reproductive senescence occurs in the Seychelles warbler [reviewed in @Hammers2015], making this system well suited to examining how telomere length predicts lifetime reproductive success.  We expect that by gaining a fuller understanding of telomere dynamics in natural populations, the fields of life-history evolution and evolutionary ecology will be greatly enhanced.

###Acknowledgements
We thank Nature Seychelles for facilitating the long-term Seychelles warbler project. The Seychelles Bureau of Standards and Department of Environment gave permission for sampling and fieldwork. Emma Barrett laid the foundations for this study, generating the original telomere qPCR protocol. We thank everyone who has helped in the field, and the Seychelles warbler research group for discussions. This work was funded by two Natural Environment Research Council (NERC) grants to DSR (NE/F02083X/1 and NE/K005502/1). LGS was also funded by an EGI Fellowship from the University of Oxford, and HLD by a NERC fellowship.

###References

