---
output:
  word_document:
    reference_docx: myref.docx
---

###Supplementary information for Spurgin *et al.* "Telomeres reveal silver spoon effects in a wild population"





```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev=c('png','pdf'))

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')


```

**Table S1** Model selection table for analysis of factors affecting early life telomere length in juvenile Seychelles warblers. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. TQ = territory quality.

```{r Table S1}

#Run the model
lm1 <- lmer(LogTL ~ Agemonths * (Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ) + (1|LayYear)+(1|PlateID)+(1|BirdID),
            data = juv,
            REML = FALSE)
lm1 <- standardize(lm1)

#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))

#Sort out row/col names
nams <- colnames(t1)

nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'


colnames(t1) <- nams

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:14)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

```



**Table S2** Model selection table for analysis of factors affecting early life telomere loss in Seychelles warbler chicks, using longitudinal data. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. TQ = territory quality.

```{r Table S2}
temp <- subset(subset(Loss,Agemonths == 1),TimeDiff>0.3)

#Run the model
lm2 <- lmer(TROC ~ Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)


#Model averaging
t1 <- as.matrix(subset(lm3, delta<=6))



#Sort out row/col names
nams <- colnames(t1)

nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'


colnames(t1) <- nams

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:7)),t1[,c('df','AICc','delta','weight')],stringsAsFactors = FALSE)
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
tableS1[,3:5] <- lapply(tableS1[,3:5],as.numeric)
kable(tableS1,digits=2)

```



**Figure S1**
```{r Figure S1}
#####################
#4A
#####################
temp <- subset(subset(Loss,Agemonths == 1),TimeDiff>0.3)

#Run the model
lm2 <- lmer(TROC ~ Tarsus + Helper + NonHelper + Sex + BodyMass  + TQ + (1|LayYear),
            data = temp,
            REML = FALSE)
lm2 <- standardize(lm2)
#Model averaging

lm3 <- dredge(lm2)
lm3 <- subset(lm3,delta <=6)
lmavg <- model.avg(lm3,fit=T)
myimp <- importance(lmavg) #Relative importance of each explanatory variable
  

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)


#Sort out row/col names
nams <- rownames(table2)
nams[nams == 'z.NonHelper'] <- 'Non helpers'
nams[nams == 'c.Sex'] <- 'Sex'
nams[nams == 'z.TQ'] <- 'TQ'
nams[nams == 'c.Helper'] <- 'Helpers'
nams[nams == 'z.BodyMass'] <- 'Body mass'
nams[nams == 'z.Agemonths'] <- 'Age'
nams[nams == 'z.LogTL'] <- 'Initial TL'
nams[nams == 'z.Tarsus'] <- 'Tarsus'
nams[nams == 'z.Agemonths:z.NonHelper'] <- 'Age x non helpers'
nams[nams == 'c.Sex:z.Agemonths'] <- 'Age x sex'
nams[nams == 'z.Agemonths:z.TQ'] <- 'Age x TQ'
nams[nams == 'c.Helper:z.Agemonths'] <- 'Age x helpers'
nams[nams == 'z.Agemonths:z.BodyMass'] <- 'Age x mass'
nams[nams == 'z.Agemonths:z.Tarsus'] <- 'Age x tarsus'

rownames(table2) <- nams
colnames(table2) <- c('Estimate','SE','2.5%','97.5%','RI')


table2 <- data.frame(table2,myx=row.names(table2))

FigS1 <- forestplot(table2)
FigS1

```