---
output: word_document
---

###Supplementary information for Spurgin *et al.* "Telomeres reveal complex silver spoon effects in a wild population"





```{r load and clean, include = F}

#Set wd one level up
setwd('..')


#Load and clean data

source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')


#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = getwd())
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 4,
               fig.height= 8,
               fig.show='asis',
               dev='png')

```


**Table S1** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler chicks. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. i = intercept.

```{r Table S1}

#Run the model
lm1 <- lmer(TLKB ~ TQspace + Insect + NonHelper + Sex + BodyMass + Tarsus + Helper + (1|LayYear),
            data = chickall,REML=FALSE)

lm1 <- standardize(lm1)


#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))
colnames(t1)[2:8] <- c('Sex', 'Body Mass','Helpers','Food av.','Non helpers','Tarsus length','TQ')

tableS1 <- data.frame(Model = apply(t1,1,getterms,c(2:8)),t1[,c('df','AICc','delta','weight')])
colnames(tableS1) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tableS1) <- NULL
tableS1$Model[tableS1$Model == ''] <- 'Null model'
kable(tableS1,digits=3)

```





**Table S2** Model selection table for analysis of factors affecting early life telomere length in Seychelles warbler fledglings and subadults. Only models that remained in the top model set for model averaging (AICc $\leq$ 6 compared to the top model) are included here. i = intercept.

```{r Table s2}

#Run the model
lm1 <- lmer(TLKB ~ TQspace + Insect + NonHelper + Sex + BodyMass + Tarsus + Helper + (1|LayYear),
            data = FlSAall,REML=FALSE)

lm1 <- standardize(lm1)


#Model averaging
lm2 <- dredge(lm1)
t1 <- as.matrix(subset(lm2, delta<=6))
colnames(t1)[2:8] <- c('Sex', 'Body Mass','Helpers','Food av.','Non helpers','Tarsus length','TQ')

tables2 <- data.frame(Model = apply(t1,1,getterms,c(2:8)),t1[,c('df','AICc','delta','weight')])
colnames(tables2) <- c('Model','df','AICc', 'Delta AICc','Weight')
rownames(tables2) <- NULL
tables2$Model[tables2$Model == ''] <- 'Null model'
kable(tables2,digits=3)

```










PAGEBREAK

```{r Figure S1,fig.width=4,fig.height=4}


temp <- dd

ddFigS1 <- ddply(dd,
                 .(Age),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS1 <- ddFigS1[complete.cases(ddFigS1),]
ddFigS1 <- subset(ddFigS1,Age<14)

FigS1 <- ggplot(ddFigS1,
       aes(x = Age,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  ylab('Telomere length (kb)') + 
  xlab('Age') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS1

```


**Figure S1** Cross-sectional telomere length (mean $\pm$ s.e.) in relation to age in Seychelles warblers.

PAGEBREAK









```{r Figure S2,fig.width=4,fig.height=4}

temp <- Loss
levels(temp$Ageclass) <- c('Chicks', 'Fledglings','Subadults')

ddFigS2 <- ddply(temp,
                 .(Ageclass),
                 summarise,
                 TROCmean = mean(TROC),
                 TROCse = se(TROC))

FigS2 <- ggplot(ddFigS2,
       aes(x = Ageclass,
           y = TROCmean)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('') + 
  theme_classic() +
  ylim(0,0.002) +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS2


```

**Figure S2** Telomere loss (mean $\pm$ s.e.) in juvenile Seychelles warblers. Telomere loss a standardised and corrected value (see methods for details) based on rates of lost between the sampling event of birds as a juvnile, and the first available sampling as an adult. Values above zero indicate telomere shortening in early life.









```{r Figure S3,fig.width=4,fig.height=4}


temp <- FlSA

ddFigS3 <- ddply(temp,
                 .(LayYear),
                 summarize,
                 TLKBmean = mean(TLKB),
                 TLKBse = se(TLKB))

ddFigS3 <- ddFigS3[complete.cases(ddFigS3),]
ddFigS3$LayYear <- as.numeric(paste(ddFigS3$LayYear))

FigS3 <- ggplot(ddFigS3,
       aes(x = LayYear,
           y = TLKBmean)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  ylab('Telomere length (kb)') + 
  xlab('Lay Year') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16))

FigS3

```


**Figure S3** Cross-sectional juvenile telomere length (mean $\pm$ s.e.) across birth years in Seychelles warblers