---
output:
  beamer_presentation:
    fig_caption: yes
    colortheme: 'dove'
---

```{r load clean, include = FALSE}
#Load and clean data
rm(list=ls())
setwd('..')
source('Scripts/SW juvenile telomeres IPC load.r')
source('Scripts/SW juvenile telomeres IPC functions.r')
source('Scripts/SW juvenile telomeres IPC clean.r')

```






## **In early life?**

\begin{center}
  \includegraphics{Pictures/SWfeedingcomp.png}
\end{center}







## **Questions**

####1. Is telomere loss greatest in the first year of life?



####2. What factors affect telomere length in early life?  


####3. What are the consequences of early life telomere dynamics?

\begin{center}
  \includegraphics{Pictures/SWfeedingcomp.png}
\end{center}











## **Is telomere loss greatest in the first year of life?**

```{r, echo=FALSE,fig.height=5,fig.width=8,fig.cap='Telomere length and age'}
par(pty='s')
errorplot(y = dd$TLKB,
          x = dd$Age,
          myylab = 'Telomere Length (kb)',
          myxlab = 'Age')
polygon(x=c(-0.2,0.81,0.81,-0.2),
        y=c(5.5,5.5,6,6),
        col='white',
        border = NA)
```










## **Is telomere loss greatest in the first year of life?**

```{r, echo=FALSE,fig.height=5,fig.width=8, fig.cap='Telomere length and age'}
par(pty='s')
errorplot(y = dd$TLKB,
          x = dd$Age,
          myylab = 'Telomere Length (kb)',
          myxlab = 'Age')
```












## **Is telomere loss greatest in the first year of life?**

```{r, echo=FALSE,fig.height=5,fig.width=8, fig.cap='Telomere length and juvenile age'}
par(pty='s')
errorplot(y = dd$TLKB,
          x = dd$Ageclass,
          myylab = 'Telomere Length (kb)',
          myxlab = 'Age Class')
```











## **Questions**

####1. Is telomere loss greatest in the first year of life?  
*Yes, but it's lost at a very early stage*

####2. What factors affect telomere length in early life?  


####3. What are the consequences of early life telomere dynamics?

\begin{center}
  \includegraphics{Pictures/SWfeedingcomp.png}
\end{center}












## **What factors affect telomere length in early life?**

####Model including only juvenile birds (*N* = `r nrow(juv)`)

```{r, eval = FALSE}
mymodel <- lm(Telomere_Length ~ Mass + 
                                Tarsus_Length + 
                                Age_Class + 
                                Sex + 
                                Territory_Quality + 
                                EPP + 
                                Helper,
              data = juveniles)
```









## **What factors affect telomere length in early life?**

####Key factors:

- Territory Quality
- Age class x Sex
- EPP

\begin{center}
  \includegraphics{Pictures/cousinandchicks.png}
  \end{center}








## **What factors affect telomere length in early life?**
```{r, echo=FALSE,fig.height=5,fig.width=8,fig.cap='Telomere length and territory quality',fig.pos='H'}
par(pty='s')
errorplot(juv$TLKB,juv$LogTQF,
          myxlab = 'Territory Quality',
          myylab = 'Telomere Length (kb)')
  
```








## **What factors affect telomere length in early life?**
```{r, echo=FALSE,fig.height=4,fig.width=8, fig.cap='Telomere length and Sex * Age'}
layout(matrix(1:2,1,2))
par(mar = c(5,5,2,0))
errorplot(y = m_juv$TLKB,
          x = m_juv$Ageclass,
          myylab = 'Telomere Length (kb)',
          myxlab = 'Age Class',
          mymain = 'Males',
          myylim = c(5,7),
          myxlim <- c(0,1))

par(mar=c(5,3,2,2))
errorplot(y = f_juv$TLKB,
          x = f_juv$Ageclass,
          myylab = '',
          myxlab = 'Age Class',
          mymain = 'Females',
          myylim = c(5,7),
          myxlim <- c(0,1))
```







## **What factors affect telomere length in early life?**
```{r, echo=FALSE,fig.height=4,fig.width=8, fig.cap='Telomere length and EPP * Age'}
layout(matrix(1:2,1,2))
par(mar = c(5,5,2,0))
errorplot(y = juvnotEPP$TLKB,
          x = juvnotEPP$Ageclass,
          myylab = 'Telomere Length (kb)',
          myxlab = 'Age Class',
          mymain = 'Within Pair',
          myylim = c(5,7),
          myxlim <- c(0,1))

par(mar=c(5,3,2,2))
errorplot(y = juvEPP$TLKB,
          x = juvEPP$Ageclass,
          myylab = '',
          myxlab = 'Age Class',
          mymain = 'Extra Pair',
          myylim = c(5,7),
          myxlim <- c(0,1))
```












## **Questions**

####1. Is telomere loss greatest in the first year of life?  
*Yes, but it's lost at a very early stage*

####2. What factors affect telomere length in early life?  
*Age, Sex, Territory Quality and EPP*

####3. What are the consequences of early life telomere dynamics?


\begin{center}
  \includegraphics{Pictures/SWfeedingcomp.png}
\end{center}




## **What are the consequences of early life telomere length?**
```{r, echo=FALSE,fig.height=4,fig.width=8, fig.cap='Telomere length and survival'}


layout(matrix(1:2,1,2))


######
#Survival to next year
######

errorplot(x=juv$SurvivedNextFig,y=juv$TLKB,
          myxlim = c(-0.5,1.5),
          myylim = c(5,6),
          myylab = 'Telomere Length (kb)',
          myxlab = 'Survived to adulthood?')

######
#Remaining Life
######

plot1(RemainingLife~TLKB,
      mydata=juv,
      myxlab = 'Telomere Length (kb)',
      myylab = 'Remaining Life',
      myylim=c(0,20))

#Run a model(for fit line)
glm1 <- zeroinfl(RemainingLife~TLKB,
      data=juv,
      dist='poisson')

#Get fitted values and fit the line
predProbs<-predict(glm1,data.frame(TLKB=seq(1, 12, length.out=100)), type="response")
lines(seq(1, 12, length.out=100), predProbs, col='red', lwd=2)

```












## **What are the consequences of early life telomere length?**
```{r, echo=FALSE,fig.height=4,fig.width=8, fig.cap='Telomere length and survival'}


layout(matrix(1:2,1,2))


######
#Survival to next year
######

errorplot(x=m_juv$SurvivedNextFig,y=m_juv$TLKB,
          myxlim = c(-0.5,1.5),
          myylim = c(4,7),
          myylab = 'Telomere Length (kb)',
          myxlab = 'Survived to adulthood?')
errorplot(x=f_juv$SurvivedNextFig,y=f_juv$TLKB,
          add=T,
          mylegend=T,
          legendlevels = c('Males','Females'))

######
#Remaining Life
######

plot1(RemainingLife~TLKB,
      mydata=juv,
      myxlab = 'Telomere Length (kb)',
      myylab = 'Remaining Life',
      myylim=c(0,20))

#Run a model(for fit line)
glmMale <- zeroinfl(RemainingLife~TLKB,
      data=m_juv,
      dist='poisson')
glmFemale <- zeroinfl(RemainingLife~TLKB,
      data=f_juv,
      dist='poisson')

#Get fitted values and fit the line
predProbs<-predict(glmMale,data.frame(TLKB=seq(1, 12, length.out=100)), type="response")
lines(seq(1, 12, length.out=100), predProbs, col='red', lwd=2)
predProbs<-predict(glmFemale,data.frame(TLKB=seq(1, 12, length.out=100)), type="response")
lines(seq(1, 12, length.out=100), predProbs, col='blue', lwd=2)

legend('topleft',
       pch='--',
       bty='n',
       col=c('red','blue'),
       legend=c('Males','Females'),
       cex=1.2)

```













## **What are the consequences of early life telomere length?**
```{r, echo=FALSE,fig.height=4,fig.width=8, fig.cap='Telomere length and survival'}


layout(matrix(1:2,1,2))


######
#Survival to next year
######

errorplot(x=juvEPP$SurvivedNextFig,y=juvEPP$TLKB,
          myxlim = c(-0.5,1.5),
          myylim = c(4.5,7.5),
          myylab = 'Telomere Length (kb)',
          myxlab = 'Survived to adulthood?')
errorplot(x=juvnotEPP$SurvivedNextFig,y=juvnotEPP$TLKB,
          add=T,
          mylegend=T,
          legendlevels = c('Extra Pair','Within Pair'))

######
#Remaining Life
######

plot1(RemainingLife~TLKB,
      mydata=juv,
      myxlab = 'Telomere Length (kb)',
      myylab = 'Remaining Life',
      myylim=c(0,20))

#Run a model(for fit line)
glmEPP <- zeroinfl(RemainingLife~TLKB,
      data=juvEPP,
      dist='poisson')
glmnotEPP <- zeroinfl(RemainingLife~TLKB,
      data=juvnotEPP,
      dist='poisson')

#Get fitted values and fit the line
predProbs<-predict(glmEPP,data.frame(TLKB=seq(1, 12, length.out=100)), type="response")
lines(seq(1, 12, length.out=100), predProbs, col='red', lwd=2)
predProbs<-predict(glmnotEPP,data.frame(TLKB=seq(1, 12, length.out=100)), type="response")
lines(seq(1, 12, length.out=100), predProbs, col='blue', lwd=2)

legend('topleft',
       pch='--',
       bty='n',
       col=c('red','blue'),
       legend=c('Extra Pair','Within Pair'),
       cex=1.2)

```









## **Questions**

####1. Is telomere loss greatest in the first year of life?  
*Yes, but it's lost at a very early stage*

####2. What factors affect telomere length in early life?  
*Age, Sex, Territory Quality and EPP*

####3. What are the consequences of early life telomere dynamics?
*Important for survival but dependent on early life conditions*

\begin{center}
  \includegraphics{Pictures/SWfeedingcomp.png}
\end{center}








## **Thanks!**

- Pat Monaghan
- Winnie Boner
- Simon Verhulst

**Seychelles warbler study group**  (*Emma Barrett*, Jan Komdeur, Terry Burke and all the other warbler researchers)

\begin{center}
  \includegraphics{Pictures/UEAlogo.png}
\end{center}

\begin{center}
  \includegraphics{Pictures/NERClogo.png}
\end{center}